Automated Formal Synthesis of Digital Controllers for State-Space Physical Plants

### 2017-01-23 cav2017/cav2017.tex
Stability bug fix Continuous time envelope proof

Whilst there is a vast literature on validation of control systems,
the subject of control synthesis has been somewhat restricted in the
properties it explores.  Traditionally, this focuses mainly on the
subject of stability %, working typically on continuous domains
\cite{DBLP:journals/corr/AbateBCCDKK16,sadabadi2016static}, or using switched
models~\cite{DBLP:conf/emsoft/RavanbakhshS16}.
Reachability tools are
also used when discretizing the state-space for both verification and
parametric synthesis~\cite{cimatti2013parameter}.

In this work, we propose the synthesis of \emph{safe} control algorithms for
LTI models taking into consideration the continuous domain of the plant, 
alongside the discrete domain of the controller and the hybrid elements that relate them.  
%% In this paper, we synthesize controllers for hybrid
%% closed-loop systems, i.e., software-implemented embedded controllers along
%% with a model of their physical environment (the plant), that are
%% \emph{safe}.
Due to the complexity of such systems, in this particular work we focus
on linear models with known  implementation features (e.g., number of bits, fixed-point
arithmetic) %\addtodo{[clarify next] with known configurations} 
and perform parametric synthesis of safe digital controllers.

@@

Program synthesisers are an ideal fit for synthesis of parametric
controllers since the semantics of programs capture effects such as FWL
precisely.  In~\cite{DBLP:conf/cdc/RavanbakhshS15}, the authors use CEGIS
for the synthesis of switching controllers for stabilizing continuous-time
plants with polynomial dynamics.  The work extends to its application on
affine systems, finding its major challenge in the hardness of solving
linear arithmetic with the state-of-the-art SMT solvers.  Since this
approach uses switching states instead of linear dynamics in the digital
controller, it entirely circumvents the FWL problem.  It is also not
suitable for the kind of control we seek to synthesize.
\textcolor{red}{Cristina: can you please revise this sentence:
Moreover, in \cite{DBLP:conf/emsoft/RavanbakhshS16} the same authors 
use a CEGIS-based approach for synthesizing continuous-time switching
controllers that guarantee reach while stay properties of a closed
loop system}. The solution is based on synthesizing control Lyapunov
functions for switched systems, that yield switching controllers with
a guaranteed minimum dwell time in each mode.

In \cite{DBLP:journals/corr/AbateBCCDKK16}, Abate et al.  focus on
synthesising stabilizing controllers for continuous plants
by exploiting advantage in bit-accurate verification of software-implemented 
digital control~\cite{Bessa16}.  While they also make use of a CEGIS-based 
technique, their approach is restricted to synthesising controllers for hybrid 
closed-loop systems that are stable and only consider transfer function models.  
In contrast, in the current paper, we consider a state-space
representation of the physical system, which allows us to synthesise controllers that
make the closed-loop system \emph{safe} as well as stable.  Moreover, 
a key aspect of the second approach we evaluate
is the fact that we integrate abstraction
refinement inside the CEGIS loop.

We use a combination of a synthesis engine with a control
verification tool that addresses the challenges presented here in the
form of FWL effects and stability measures for LTI SISO controllers.
We take the former from~\cite{DBLP:conf/lpar/DavidKL15} and the latter
from~\cite{Monteiro16}%\todo{This paper seems to be for transfer functions, not state space,
%is there a paper that does state space stability and safety?}
, while enhancing the procedure by evaluating the
quantization effects of the Hardware interfaces (ADC/DAC) to obtain an
accurate discrete-time FWL representation of the continuous dynamics.

@@

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\subsection{Overapproximating continuous space dynamics using discrete models}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
Let us recall equation \ref{eq:discretize}. The discrete representation of the system dynamics discretized with sample time $T_1$ is ruled by the formula
$\vec{A}_1 = e^{\vec{A} T_1}$. From matrix theory, we have 
\begin{equation}
\vec{A}_1 = \mat{S}_1\mat{J}_1\mat{S}_1^{-1}= \mat{S}_1e^{\mat{J} T_1}\mat{S}_1^{-1}.
\end{equation}
Let $\mat{J}$ be a diagonal matrix, such that ${\lambda_1}_i=e^{\lambda_i T_1}, \forall \lambda \in \mat{J}$.
Let us now take a different sample rate $T_2$, such that ${\lambda_2}_i=e^{\lambda_i T_2}, \forall \lambda \in \mat{J}$. We can then say that 
\begin{equation}
{\lambda_2}_i=e^{\lambda_i T_1 T_2/T_1}={\lambda_1}_i^{T_2/T_1} \Rightarrow A_2=A_1^{\frac{T_2}{T_1}}
\end{equation}
We have also shown in the previous section that $x_k=A^nx_0$ is a witness of $x(t): \dot{x}(t)=Ax(t)$ at time $kT_s$. Applying this notion to multiple sample times, we can state that $x_{k_i}=A_i^nx_0$ are a set of witnesses at times $kT_i$, and by extension the set $\{x=A_i^nx_0 : \forall n \in \mathbb{R} \}$ witnesses all states visited by the continuous system in time $t \in [0,\infty)$. This applies to any arbitrary sample time $T_i$.

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\subsection{Modelling quantization as noise}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++

>>>>

Whilst there is a vast literature on validation of control systems,
the subject of control synthesis has been somewhat restricted in the
properties it explores.  Traditionally, this focuses mainly on stability %, working typically on continuous domains
\cite{DBLP:journals/corr/AbateBCCDKK16,sadabadi2016static}, or using switched
models~\cite{DBLP:conf/emsoft/RavanbakhshS16}.
Reachability tools are
also used when discretizing the state-space for verification and
parametric synthesis~\cite{cimatti2013parameter}.

In this work, we propose the synthesis of \emph{safe} control algorithms for
LTI models taking into consideration the continuous domain of the plant, 
alongside the discrete domain of the controller and the hybrid elements that relate them.  
%% In this paper, we synthesize controllers for hybrid
%% closed-loop systems, i.e., software-implemented embedded controllers along
%% with a model of their physical environment (the plant), that are
%% \emph{safe}.
Due to the complexity of such systems, in this particular work we focus
on linear models with known  implementation features (e.g., number of bits, fixed-point
arithmetic) %\addtodo{[clarify next] with known configurations} 
and perform %parametric 
synthesis of safe digital controllers.

@@

Program synthesisers are an ideal fit for synthesis of %parametric
controllers since the semantics of programs capture effects such as FWL
precisely.  In~\cite{DBLP:conf/cdc/RavanbakhshS15}, the authors use CEGIS
for the synthesis of switching controllers for stabilizing continuous-time
plants with polynomial dynamics.  The work extends to its application on
affine systems, finding its major challenge in the hardness of solving
linear arithmetic with the state-of-the-art SMT solvers.  Since this
approach uses switching states instead of linear dynamics in the digital
controller, it entirely circumvents the FWL problem.  It is also not
suitable for the kind of control we seek to synthesize.
\textcolor{red}{Cristina: can you please revise this sentence:
Moreover, in \cite{DBLP:conf/emsoft/RavanbakhshS16} the same authors 
use a CEGIS-based approach for synthesizing continuous-time switching
controllers that guarantee reach while stay properties of a closed
loop system}. The solution is based on synthesizing control Lyapunov
functions for switched systems, that yield switching controllers with
a guaranteed minimum dwell time in each mode.

In \cite{DBLP:journals/corr/AbateBCCDKK16}, Abate et al.  focus on
synthesising stabilizing controllers for continuous plants
by exploiting advantage in bit-accurate verification of software-implemented 
digital control~\cite{Bessa16}.  While they also make use of a CEGIS-based 
technique, their approach is restricted to synthesising controllers for hybrid 
closed-loop systems that are stable and only consider transfer function models.  
In contrast, in the current paper, we consider a state-space
representation of the physical system, which allows us to synthesise controllers that
make the closed-loop system \emph{safe} as well as stable.  Moreover, 
a key aspect of the second approach we evaluate
is that we integrate abstraction
refinement inside the CEGIS loop.

The tool Pessoa~\cite{mazo2010pessoa} is a MatLab toolbox that synthesises correct-by-design
embedded control software. It is based on the principle of describing a physical system
as an equivalent finite-state machine and computing reachability over the finite-state 
machine. Pessoa can synthesise safe embedded controller software for a range of properties based on a safety specification over state
values, as well as specification on characteristic functions of the controller. 
However, Pessoa does not have native inclusion of the ADC/DAC quantization error needed for
synthesising a fixed point digital controller.

We use a combination of a synthesis engine with a control
verification tool that addresses the challenges presented here in the
form of FWL effects and stability measures for LTI SISO controllers.

We take the former from~\cite{DBLP:conf/lpar/DavidKL15}. The verification engine, for
the first approach we present, is 
based on~\cite{Monteiro16}, but 
enhanced to include evaluation of quantization effects and safety properties
over the state space. In the second approach, we use a verification engine
based on XXXXXXX
\todo{something about the acceleration verification engine}

@@

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\subsection{Overapproximating continuous space dynamics using discrete models}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
Let us recall equation \ref{eq:discretize}. The discrete representation of the system dynamics discretized with sample time $T_1$ is ruled by the formula
$\vec{A}_1 = e^{\vec{A} T_1}$. From matrix theory, we have 
\begin{equation}
\vec{A}_1 = \mat{S}_1\mat{J}_1\mat{S}_1^{-1}= \mat{S}_1e^{\mat{J} T_1}\mat{S}_1^{-1}.
\end{equation}
Let $\mat{J}$ be a diagonal matrix, such that ${\lambda_1}_i=e^{\lambda_i T_1}, \forall \lambda \in \mat{J}$.
Let us now take a different sample rate $T_2$, such that ${\lambda_2}_i=e^{\lambda_i T_2}, \forall \lambda \in \mat{J}$. We can then say that 
\begin{equation}
{\lambda_2}_i=e^{\lambda_i T_1 T_2/T_1}={\lambda_1}_i^{T_2/T_1} \Rightarrow A_2=A_1^{\frac{T_2}{T_1}}
\end{equation}
We have also shown in the previous section that $x_k=A^nx_0$ is a witness of $x(t): \dot{x}(t)=Ax(t)$ at time $kT_s$. Applying this notion to multiple sample times, we can state that $x_{k_i}=A_i^nx_0$ are a set of witnesses at times $kT_i$, and by extension the set $\{x=A_i^nx_0 : \forall n \in \mathbb{R} \}$ witnesses all states visited by the continuous system in time $t \in [0,\infty)$. This applies to any arbitrary sample time $T_i$.
Since the controller output is fed to a Zero Order Holder, the progression of the system between times $kT_s$ and $(k+1)T_s$ is $\mat{A}^n\vec{x}_k+\mat{B}\vec{u}_k : n \in (0,1)$. We therefore obtain a reach tube for the continuous dynamics of a time sampled system to be contained within the abstract reach tube $\mathcal{A}^1\mathcal{M}X_0+\mathcal{BK}^1\mathcal{M}X_0=\mathcal{M}X_0$
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\subsection{Modelling quantization as noise}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
